---
title: "Prepare data for GFBLUP"
author: "S. Turner"
date: "January 31, 2018"
output: html_document
---
## Summary
This script prepares the genotype matrix (W) and the genomic relationship matrix (G) to run GBLUP and GFBLUP

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir="~/Documents/AA-GenomicPrediction/")
library(qgg)
# source("https://bioconductor.org/biocLite.R")
# biocLite("snpStats")
library(snpStats)
```

## Load and prepare data

### Phenotype and co-variable data
Note: no co-variable info for this data (previously accounted for)
Removes individuals with missing phenotypid data and writes out clean file ('phenotypes.csv')
```{r phenotypes}
phenotypes <- read.csv("./data/input_files/AfterBlupsAbsLevels04222014.csv")
phenotypes$Taxa <- paste0("A", phenotypes$Taxa) # append 'A' to taxa names
phenotypes <- phenotypes[-78,] # remove individual with no genotypic data
ratios <- read.csv("~/Documents/AA-GenomicPrediction/data/input_files/Blup-transformedforGWAS.csv", header=TRUE)
ratios$Taxa <- paste0("A", ratios$Taxa)
phenotypes <- merge(phenotypes, ratios, by = "Taxa")
phenotypes <- replace(phenotypes, is.na(phenotypes), NA)
write.csv(phenotypes, "./data/input_files/phenotypes.csv", row.names=FALSE)
```

### Centered and scaled genotype matrix (W)
Note: Use Wlist later to extract genomic relationship matrix (G)
```{r genotype matrix}
# Construct the genotype matrix (W)
W_full <- read.plink(bed="./data/GS_subset/plinkGeneOmeSubset.bed", bim="./data/GS_subset/plinkGeneOmeSubset.bim", fam="./data/GS_subset/plinkGeneOmeSubset.fam", na.strings=c("-9"), sep=".")

W <- W_full$genotypes
rm(W_full)
rownames(W) <- paste("A", rownames(W), sep="")
W <- as.matrix(as.data.frame(W))
class(W) <- "numeric"
W <- scale(W, center=TRUE, scale=TRUE)
dim(W)
W[1:5,1:5]

W <- W[rownames(W)%in%phenotypes$Taxa,] # remove individuals with missing phenotypic data
dim(W)

# extract names of SNPs with missing data and remove from W
m_missing <- colnames(W)[colSums(is.na(W)) > 0] # markers with missing data in full set
W <- W[,!colSums(is.na(W)) > 0]
dim(W)

write.matrix(W, "./data/input_files/W_mat.txt")

## EDIT - this no longer works
## this appears to be different as of 14 August 2018
Glist <- prepG(bedfiles="./data/GS_subset/plinkGeneOmeSubset.bed",
               bimfiles="./data/GS_subset/plinkGeneOmeSubset.bim",
               famfiles="./data/GS_subset/plinkGeneOmeSubset.fam", study="AA-gfblup",
               fnRAW="./data/GS_subset/plinkGeneOmeSubset.out")

Glist$ids <- paste("A", Glist$ids, sep="")

W2 <- getW(Glist)
```

### Compute G from W
G is the additive genomic relationship matrix, constructed using all genetic markers as G=WW'/m, where W is the centered and scaled genotype matrix and m is the total number of markers 
# not sure if necessary... 
```{r compute G from W}
# remove individuals with missing phenotype
L <- phenotypes$Taxa
W <- get
# G <- computeG(Glist = W)
# rm(Wlist)
# G <- G[rownames(G)%in%phenotypes$Taxa, colnames(G)%in%phenotypes$Taxa] # remove individuals with missing data
# G[L,L]
# dim(G)
# 
# write.matrix(G, "./data/input_files/G_mat.txt")
```

### Markers for amino acid subset
Can do this more easily in plink.. but just need to do it once so oh well 
```{r marker data for amino acid subset}
# get SNP names for AAS
W_aas <- read.plink(bed="./data/AAS_subset/plinkAASubset.bed", bim="./data/AAS_subset/plinkAASubset.bim", fam="./data/AAS_subset/plinkAASubset.fam", na.strings=c("-9"), sep=".")

# extract marker names for AAS
m_aas <- colnames(W_aas$genotypes@.Data)

rm(W_aas)

write.table(m_aas, "./data/input_files/aas_snplist.txt")
```

### Markers for control subsets
Note: removed markers with missing data

```{r marker data for control subsets}
# use list of markers with missing values to remove missing data in control subsets
for (i in 1:1000){
  path <- paste("./data/control_markers/plink_control_snplist_", i, ".txt", sep="")
  m_control <- read.table(path)
  m_control <- as.character(m_control[,1])
  m_control <- grep(paste(m_missing, collapse="|"), m_control, value=TRUE, invert=TRUE)
  print(length(m_control))
  write.table(m_control, paste("./data/control_markers/plink_control_snplist_", i, ".txt", sep=""))
}

rm(i, path, m_control)

```

```{r save rdata}
save.image(file="./data/input_files/gfblup_input.Rdata")
```